---
title: "Tidy your Data !"
excerpt: "Resources for Data Management in R"
collection: R
---

```{r, include=FALSE}
library(tidyverse)
```

 
## Tibbles vs data.frame

Tibbles have a refined print method that shows only the first 10 rows and each column reports its type, function borrowed from str() :

```{r}
as_tibble(iris)
```

You can also control the amount of rows (n) being printed. In this case, we specify width = Inf, which will print all columns

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```

## Subsetting

**$ and [[**  can extract by name or position while **$** can only extracts by name:

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)
#Extract by name
df$x
df[["x"]]
#Extract by position
df[[1]]
```

To use these in a pipe, we need to use the special placeholder . :
```{r}
df %>% 
    .$x
df %>% 
    .[["x"]]
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

 